---
title: "Assignment 5"
author: "Hemanth Allamaneni"
date: "2023-10-8"
categories: [code, graphs]
---

### Histogram

```{r}
# Generate a random dataset
set.seed(123)
data <- rnorm(1000)

# Create a histogram
hist(data, col = "skyblue", main = "Histogram Example", xlab = "Values", ylab = "Frequency")
```

### Bar charts, vertical and horizontal

```{r}
# Generate a sample dataset
categories <- c("Category A", "Category B", "Category C", "Category D")
values <- c(25, 40, 30, 20)

# Create a vertical bar chart
barplot(values, names.arg = categories, col = "skyblue", main = "Vertical Bar Chart", xlab = "Categories", ylab = "Values")

# Create a horizontal bar chart
barplot(values, names.arg = categories, col = "lightcoral", horiz = TRUE, main = "Horizontal Bar Chart", xlab = "Values", ylab = "Categories")
```

### Pie Chart

```{r}
# Generate a sample dataset
categories <- c("Category A", "Category B", "Category C", "Category D")
values <- c(25, 40, 30, 20)

# Create a pie chart
pie(values, labels = categories, col = rainbow(length(categories)), main = "Pie Chart")
```

### Box plot

```{r}
# Generate a sample dataset (replace this with your own data)
set.seed(123)
data <- list(group1 = rnorm(100, mean = 0, sd = 1),
             group2 = rnorm(100, mean = 2, sd = 1),
             group3 = rnorm(100, mean = 1, sd = 1))

# Create a boxplot
boxplot(data, col = c("skyblue", "lightcoral", "lightgreen"), main = "Boxplot Example", xlab = "Groups", ylab = "Values")
```

### Scatter plot

```{r}
# Generate a sample dataset (replace this with your own data)
set.seed(123)
x <- rnorm(100)
y <- 2*x + rnorm(100)

# Create a scatterplot
plot(x, y, col = "darkorange", main = "Scatterplot Example", xlab = "X-axis", ylab = "Y-axis")

```

## GGPlot Charts

### Histogram

```{r}
# Load the ggplot2 package
library(ggplot2)

# Generate a random dataset (replace this with your own data)
set.seed(123)
data <- rnorm(1000)

# Create a histogram using ggplot2
ggplot(data.frame(x = data), aes(x)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  ggtitle("Histogram Example") +
  xlab("Values") +
  ylab("Frequency")
```

### Bar chart

### Vertical

```{r}
# Generate a sample dataset (replace this with your own data)
categories <- c("Category A", "Category B", "Category C", "Category D")
values <- c(25, 40, 30, 20)
data <- data.frame(categories, values)

# Create a vertical bar chart using ggplot2
ggplot(data, aes(x = categories, y = values, fill = categories)) +
  geom_bar(stat = "identity", color = "black") +
  ggtitle("Vertical Bar Chart") +
  xlab("Categories") +
  ylab("Values")
```

### Horizontal

```{r}
# Create a horizontal bar chart using ggplot2
ggplot(data, aes(x = categories, y = values, fill = categories)) +
  geom_col() +
  ggtitle("Horizontal Bar Chart") +
  xlab("Categories") +
  ylab("Values") +
  coord_flip()  # Flipping coordinates for a horizontal chart
```

### Pie Chart

```{r}
# Create a pie chart using ggplot2
ggplot(data, aes(x = "", y = values, fill = categories)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  ggtitle("Pie Chart") +
  coord_polar("y")
```

### Box plot

```{r}
# Generate a sample dataset (replace this with your own data)
set.seed(123)
data <- data.frame(
  group = rep(c("Group 1", "Group 2", "Group 3"), each = 100),
  values = c(rnorm(100, mean = 0, sd = 1), rnorm(100, mean = 2, sd = 1), rnorm(100, mean = 1, sd = 1))
)

# Create a boxplot using ggplot2
ggplot(data, aes(x = group, y = values, fill = group)) +
  geom_boxplot() +
  ggtitle("Boxplot Example") +
  xlab("Groups") +
  ylab("Values")
```

### Scatter plot

```{r}
# Generate a sample dataset (replace this with your own data)
set.seed(123)
data <- data.frame(
  x = rnorm(100),
  y = 2 * rnorm(100) + rnorm(100)
)

# Create a scatterplot using ggplot2
ggplot(data, aes(x = x, y = y)) +
  geom_point(color = "darkorange") +
  ggtitle("Scatterplot Example") +
  xlab("X-axis") +
  ylab("Y-axis")
```

# File Format differences:

### The choice of file format for exporting graphics depends on the specific needs of the end user. PDF is ideal for high-quality printing and documents with multiple pages.

### While JPG is suitable for web graphics and photographs with millions of colors. SVG, as a scalable vector format, is optimal for web graphics and interactive charts. TIFF, known for loss less compression, is preferred in professional printing and high-quality images, whereas BMP, a simple uncompressed format, is commonly used for basic graphics in Windows applications. PNG, with loss less compression and support for transparency, is well-suited for web graphics and images requiring sharp edges.

### Each format has its unique advantages, catering to different use cases and preferences.
